---
title: "Analysis_gaze_probe_period"
author: "Dongyu Gong"
date: "23 May 2024"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup2}
set.seed(2024)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
rm(list = ls())
```

```{r library,include=TRUE,warning=FALSE,message=FALSE}
library(tidyr)
library(permuco)
library(ggplot2)
library(zoo)
library(readxl)
library(dplyr)
library(Rmisc)
library(smoothie)
library(svglite)
```

```{r fileread, results='hide'}

files <- list.files(path='Output/csv',pattern = ".csv", recursive =F) 
# load the  data
et_dat <- NULL
for(file in files){
  print(file)
  tmp <- data.table::fread(paste('Output/csv/',file,sep = ''), sep = ",", header = T)
  et_dat <- rbind(et_dat, tmp)
  tmp<-NULL
}

```

```{r}
## set values
timepersample = 1000/1000 # 1 sample = 1ms
baseline = 250 # samples, i.e. 250ms
probe_start = 1000 # probe onset
probe_end = 2000
smoothwindow = 25 # samples,i.e. 25ms
```

# X position baseline
```{r}
dat_baseline <- aggregate(meanxp ~participant+block, data=et_dat[(et_dat$cueFrame>-baseline & et_dat$cueFrame<=0),], mean)
hist(dat_baseline$meanxp,100)
names(dat_baseline)[names(dat_baseline) == "meanxp"] <- "precue_baselineX"
et_dat <- inner_join(et_dat,dat_baseline,by = c("participant","block"))
et_dat$meanxp_baselined <- et_dat$meanxp - et_dat$precue_baselineX
hist(et_dat$meanxp_baselined,100)
```

# Y position baseline
```{r}
dat_baseline <- aggregate(meanyp ~ participant+block, data=et_dat[(et_dat$cueFrame>-baseline & et_dat$cueFrame<=0),], mean)
hist(dat_baseline$meanyp,100)
names(dat_baseline)[names(dat_baseline) == "meanyp"] <- "precue_baselineY"
et_dat <- inner_join(et_dat,dat_baseline,by = c("participant","block"))
et_dat$meanyp_baselined <- et_dat$meanyp - et_dat$precue_baselineY
hist(et_dat$meanyp_baselined,100)
```

# X data after excluding the outliers
```{r}
hist(et_dat$meanxp_baselined[abs(et_dat$meanxp_baselined) < 960],100)
excl <- et_dat[abs(et_dat$meanxp_baselined) > 960,]
summary(excl$meanxp_baselined)
excl$excluded <- 1
excl_blocks <- aggregate(excluded~participant+block,excl,mean)
(avgexcl <- aggregate(excluded~participant, data=excl_blocks,sum))
avgexcl$prop <- avgexcl$excluded/600
print(paste("Avg excluded trials per participant:",(mean(avgexcl$prop))*100,"%",sep=" "))
(summarySE(data=avgexcl,measurevar = "prop"))
print(paste("Overall excluded trials:",(nrow(excl_blocks)/(600*44)*100),"%",sep=" "))
xdat <- dplyr::left_join(et_dat,excl_blocks, by = c("participant","block"))
rm(et_dat)
xdat <- xdat[is.na(xdat$excluded),]
```

# Y data after excluding the outliers
```{r}
hist(et_dat$meanyp_baselined[abs(et_dat$meanyp_baselined) < 540],100)
# have to discard data for some participants for whom 50% of the trials eye moved beyond half the distance between the fixation and the centre of the memory items
# truncated_et_dat <- et_dat[!is.element(et_dat$participant,c('32_JO','22_YW','25_YC','35_KS','50_QY','46_DP')),]
excl <- et_dat[abs(et_dat$meanyp_baselined) > 540,]
summary(excl$meanyp_baselined)
excl$excluded <- 1
excl_blocks <- aggregate(excluded~participant+block,excl,mean)
avgexcl <- aggregate(excluded~participant, data=excl_blocks,sum)
avgexcl$prop <- avgexcl$excluded/600
(avgexcl[order(avgexcl$prop,decreasing = T),])
print(paste("Avg excluded trials per participant:",(mean(avgexcl$prop))*100,"%",sep=" "))
(summarySE(data=avgexcl,measurevar = "prop"))
print(paste("Overall excluded y data:",(nrow(excl_blocks)/(600*44)*100),"%",sep=" "))
ydat <- dplyr::left_join(et_dat,excl_blocks, by = c("participant","block"))
rm(et_dat)
ydat <- ydat[is.na(ydat$excluded),]
```

# Smooth and cut X data
```{r}
xdat <- xdat[xdat$cueFrame>probe_start & xdat$cueFrame < probe_end,]
xdat$meanxp_baselined <- TTR::SMA(xdat$meanxp_baselined,smoothwindow)
```

# Smooth and cut Y data
```{r}
ydat <- ydat[ydat$cueFrame>probe_start & ydat$cueFrame < probe_end,]
ydat$meanyp_baselined <- TTR::SMA(ydat$meanyp_baselined,smoothwindow)
```

# aggregate for probe period: X
```{r}

aggX <- data.table::setDT(xdat)[,lapply(.SD, mean),
                                by=.(participant, cue_dir_lr,cue_category, cueFrame), 
                                .SDcols=c("meanxp_baselined")]

# deal with missing data
aggX$meanxp_baselined[aggX$participant=='03_SL' & aggX$cue_category=='LTM'] = zoo::na.approx(aggX$meanxp_baselined[aggX$participant=='03_SL' & aggX$cue_category=='LTM'], na.rm = FALSE, rule = 2)
```

# aggregate for probe period: Y
```{r}
aggY <- data.table::setDT(ydat)[,lapply(.SD, mean),
                                by=.(participant, cue_dir_tb,cue_category, cueFrame), 
                                .SDcols=c("meanyp_baselined")]
# deal with missing data
aggY$meanyp_baselined[aggY$participant=='03_SL' & aggY$cue_category=='LTM'] = zoo::na.approx(aggY$meanyp_baselined[aggY$participant=='03_SL' & aggY$cue_category=='LTM'], na.rm = FALSE, rule = 2)
```

# Gaze bias: left and right: X
```{r}
# aggX <- aggX[!is.na(aggX$meanxp_baselined),]
table_lr<-summarySEwithin(data=aggX, measurevar = "meanxp_baselined", withinvars = c("cue_dir_lr","cue_category","cueFrame"),idvar = "participant",na.rm = TRUE)
table_lr$time <- as.numeric(table_lr$cueFrame)

table_lr$cue_category <- factor(table_lr$cue_category, levels=c('Neutral','WM', 'LTM'))
table_lr$cue_dir_lr <- factor(table_lr$cue_dir_lr, levels=c('neutral','left', 'right'))

table_lr <- table_lr[table_lr$cue_category!='Neutral',]

ggplot(table_lr, aes(x = time, y = meanxp_baselined)) + 
    labs(y = "Offset (pixels)", x = "Time relative to probe onset (ms)", color="Cue direction") + 
    geom_hline(yintercept=0, linetype="dashed",size=1, alpha=1) +
    geom_line(aes(color = cue_dir_lr),alpha = 1,size=1) +
    geom_ribbon(aes(ymax=meanxp_baselined+se, ymin=meanxp_baselined-se, fill=cue_dir_lr), alpha = 0.2) +
    theme_set(theme_gray(base_size = 20)) +
    scale_color_manual(labels=c("left","right"),
                       values = c("blue", "red")) + 
    scale_fill_manual(labels=c("left","right"),
                      values = c("blue", "red")) + 
    # scale_y_continuous(breaks = seq(-4, 2, by = 2)) +
    guides(fill="none") +
    theme(plot.title = element_text(hjust = 0.5),         
          panel.background = element_blank(),
          legend.position= 'bottom', 
          axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"),
          strip.background = element_blank()) +
    facet_grid(~cue_category)
ggsave("figures/bias_RL.svg", device = "svg",width = 25, height = 15, units = "cm")
```

# Gaze bias: top and bottom: Y
```{r}
table_tb<-summarySEwithin(data=aggY, measurevar = "meanyp_baselined", withinvars = c("cue_dir_tb","cue_category","cueFrame"),idvar = "participant")
table_tb$time <- as.numeric(table_tb$cueFrame)

table_tb$cue_category <- factor(table_tb$cue_category, levels=c('Neutral','WM', 'LTM'))
table_tb$cue_dir_tb <- factor(table_tb$cue_dir_tb, levels=c('neutral','bottom', 'top'))

table_tb <- table_tb[table_tb$cue_category!='Neutral',]

ggplot(table_tb, aes(x = time, y = -meanyp_baselined)) + 
    labs(y = "Offset (pixels)", x = "Time relative to probe onset (ms)", color="Cue direction") + 
    geom_hline(yintercept=0, linetype="dashed",size=1, alpha=1) +
    geom_line(aes(color = cue_dir_tb),alpha = 1,size=1) +
    geom_ribbon(aes(ymax=-meanyp_baselined+se, ymin=-meanyp_baselined-se, fill=cue_dir_tb), alpha = 0.2)+
    theme_set(theme_gray(base_size = 20)) +
    scale_color_manual(labels=c("bottom","top"),
                       values = c("blue", "red")) + 
    scale_fill_manual(labels=c("bottom","top"),
                      values = c("blue", "red")) + 
    # scale_y_continuous(breaks = c(-4,-2,0,2),limits = c(-4,3)) +
    guides(fill="none") +
    theme(plot.title = element_text(hjust = 0.5),         
          panel.background = element_blank(),
          legend.position= 'bottom', 
          axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"),
          strip.background = element_blank()) + 
    facet_grid(~cue_category)

ggsave("figures/bias_TB.svg", device = "svg", width = 25, height = 15, units = "cm")
```

# Color settings for LTM and WM
```{r}
cols <- c("LTM" = rgb(138/255,118/255,190/255), "WM" = rgb(65/255,143/255,111/255))#(0.9882, 0.5529, 0.3843),(0.4000, 0.7608, 0.6471)
```

# Towardness: Left and right: X
```{r}
towardR <- aggX[aggX$cue_dir_lr == "right",]
towardL <- aggX[aggX$cue_dir_lr == "left",]
towardN <- aggX[aggX$cue_dir_lr == "neutral",]
toward <- full_join(towardR,towardL, by =c("participant","cue_category","cueFrame"))
# toward <- toward[!is.na(toward$meanxp_baselined.y),]
toward$meanxp <- (toward$meanxp_baselined.x-toward$meanxp_baselined.y)/2
table_lr_t <-summarySEwithin(data=toward, measurevar = "meanxp",withinvars = c("cue_category","cueFrame"), idvar = "participant",na.rm = TRUE)
table_lr_t$time <- as.numeric(as.character(table_lr_t$cueFrame))*timepersample

table_lr_t$cue_category <- factor(table_lr_t$cue_category, levels=c('WM', 'LTM'))

(towardnessX <- ggplot(table_lr_t, aes(x = time-1000, y = meanxp, fill=cue_category)) + 
    labs(y = "Towardness (pixels)", x = "Time relative to probe onset (ms)", color="Cue category") + 
    geom_hline(yintercept=0, linetype="dashed",size=1, alpha=1) +
    geom_vline(xintercept=0, linetype="dashed",size=0.2, alpha=0.5) +
    geom_line(size=1, aes(colour=cue_category)) +
    geom_ribbon(aes(ymax=meanxp+se, ymin=meanxp-se), alpha = 0.2)+
    scale_color_manual(labels=c("WM","LTM"),
                       values = cols,
                       aesthetics = c("colour", "fill")) + 
    # scale_y_continuous(breaks = c(0,1,2)) + #seq(1, 2, by = 1)
    guides(fill="none") +
    theme_set(theme_gray(base_size = 20)) +
    theme(plot.title = element_text(hjust = 0.5),         
          panel.background = element_blank(),
          legend.position = "bottom",
          axis.line = element_line(colour = "black", size = 0.5, linetype = "solid")
          )
 )
```

# cluster-based permutation test - X data
```{r}
#LTM
toward_LTM <- toward[toward$cue_category=="LTM",] %>% dplyr::select(participant,cueFrame, meanxp, cue_category)
eyeX <- pivot_wider(toward_LTM, names_from = cueFrame, values_from = meanxp)

randsample <- NULL
for (i in 1:length(unique(eyeX$participant))){
  r<-rnorm(n = length(table_lr_t$sd[table_lr_t$cue_category=="LTM"]) , mean = 0, sd = mean(table_lr_t$sd[table_lr_t$cue_category=="LTM"]))
  randsample <- c(randsample,r)
}
randsample <- data.frame(matrix(randsample,nrow = length(unique(eyeX$participant)), ncol = length(table_lr_t$sd[table_lr_t$cue_category=="LTM"]), byrow = T))
rand <- data.frame(participant=unique(eyeX$participant), cue_category = "random")
rand <- cbind(rand,randsample)
colnames(rand)<-colnames(eyeX)
combinedX <- rbind(eyeX,rand)

## perform cbpt
eyeClust <-permuco::clusterlm(combinedX[,(3:ncol(combinedX))] ~ cue_category + Error(participant/cue_category), 
                               data = combinedX[,-(3:ncol(combinedX))],multcomp ="clustermass", 
                               np = 10000)
plot(eyeClust, nbbaselinepts = 65, nbptsperunit = .90)
summary(eyeClust)
sig_LTM <- (which(summary(eyeClust,table_type = "full")$cue_category$`P(>) clustermass`<0.05))*timepersample

# WM
toward_WM <- toward[toward$cue_category=="WM",] %>% dplyr::select(participant,cueFrame, meanxp, cue_category)
eyeX <- toward_WM %>% pivot_wider(names_from = cueFrame, values_from = meanxp)

randsample <- NULL
for (i in 1:length(unique(eyeX$participant))){
  r<-rnorm(n = length(table_lr_t$sd[table_lr_t$cue_category=="WM"]), mean = 0, sd = mean(table_lr_t$sd[table_lr_t$cue_category=="WM"]))
  randsample <- c(randsample,r)
}
randsample <- data.frame(matrix(randsample,nrow = length(unique(eyeX$participant)), ncol = length(table_lr_t$sd[table_lr_t$cue_category=="WM"]),byrow = T))
rand <- data.frame(participant=unique(eyeX$participant), cue_category = "random")
rand <- cbind(rand,randsample)
colnames(rand)<-colnames(eyeX)
combinedX <- rbind(eyeX,rand)

## perform cbpt
eyeClust <-permuco::clusterlm(combinedX[,(3:ncol(combinedX))] ~ cue_category + Error(participant/cue_category), 
                               data = combinedX[,-(3:ncol(combinedX))],multcomp ="clustermass", 
                               np = 10000)
plot(eyeClust, nbbaselinepts = 65, nbptsperunit = .90)
summary(eyeClust)
sig_WM <- (which(summary(eyeClust,table_type = "full")$cue_category$`P(>) clustermass`<0.05))*timepersample

# Comparing two
toward_WM_LTM <- toward %>% dplyr::select(participant,cueFrame, meanxp, cue_category)
eyeX <- toward_WM_LTM %>% pivot_wider(names_from = cueFrame, values_from = meanxp)
eyeX <- eyeX[order(eyeX$cue_category),]

# perform cbpt
eyeClust <-permuco::clusterlm(eyeX[,(3:ncol(eyeX))] ~ cue_category + Error(participant/cue_category), 
                                data = eyeX[,-(3:ncol(eyeX))],multcomp ="clustermass", 
                                np = 10000)
plot(eyeClust, nbbaselinepts = baseline, nbptsperunit = .90)
summary(eyeClust)
sig_LTM_WM <- (which(summary(eyeClust,table_type = "full")$cue_category$`P(>) clustermass`<0.05))*timepersample

#Draw the figure
(towardness_sigX <-towardnessX + annotate("point", x = sig_LTM_WM , y =-3, colour = "black", size=1) + 
    annotate("point", x = sig_WM , y =-2, colour = rgb(65/255,143/255,111/255),size=1)+
    annotate("point", x = sig_LTM , y =-1, colour = rgb(138/255,118/255,190/255),size=1))

ggsave("figures/towardness_RL.svg", device = "svg",width = 25, height = 15, units = "cm")
```

# towardness: Top and bottom: Y data
```{r}

towardT <- aggY[aggY$cue_dir_tb == "top",]
towardB <- aggY[aggY$cue_dir_tb == "bottom",]
toward <- full_join(towardT,towardB, by =c("participant","cue_category","cueFrame"))

toward$meanyp <- (toward$meanyp_baselined.y-toward$meanyp_baselined.x)/2
toward <- toward[!is.na(toward$meanyp),]
table_tb_t <-summarySEwithin(data=toward, measurevar = "meanyp",withinvars = c("cue_category","cueFrame"), idvar = "participant")
table_tb_t$time <- as.numeric(table_tb_t$cueFrame)

table_tb_t$cue_category <- factor(table_tb_t$cue_category, levels=c('WM', 'LTM'))

(towardnessY <- ggplot(table_tb_t, aes(x = time, y = meanyp, fill=cue_category)) + 
    labs(y = "Towardness (pixels)", x = "Time relative to cue onset (ms)", color="Cue category") + 
    geom_hline(yintercept=0, linetype="dashed",size=1, alpha=1) +
    geom_vline(xintercept=0, linetype="dashed",size=0.2, alpha=0.5) +
    geom_line(size=1, aes(colour=cue_category)) +
    geom_ribbon(aes(ymax=meanyp+se, ymin=meanyp-se), alpha = 0.2)+
    scale_color_manual(labels=c("WM","LTM"),
                       values = cols,
                       aesthetics = c("colour", "fill")) + 
    # scale_y_continuous(breaks = c(0,1,2),limits = c(-1,2)) +
    guides(fill="none") +
    theme_set(theme_gray(base_size = 20))+
    theme(plot.title = element_text(hjust = 0.5),         
          panel.background = element_blank(),
          legend.position = "bottom",
          axis.line = element_line(colour = "black", size = 0.5, linetype = "solid")
          )
  )

```

# cluster-based permutation test- Y data

```{r}
#LTM
toward_LTM <- toward[toward$cue_category=="LTM",] %>% dplyr::select(participant,cueFrame, meanyp, cue_category)
eyeY <- pivot_wider(toward_LTM, names_from = cueFrame, values_from = meanyp)

randsample <- NULL
for (i in 1:length(unique(eyeY$participant))){
  r<-rnorm(n = length(table_tb_t$sd[table_tb_t$cue_category=="LTM"]) , mean = 0, sd = mean(table_tb_t$sd[table_tb_t$cue_category=="LTM"]))
  randsample <- c(randsample,r)
}
randsample <- data.frame(matrix(randsample,nrow = length(unique(eyeY$participant)), ncol = length(table_tb_t$sd[table_tb_t$cue_category=="LTM"]),byrow = T))
rand <- data.frame(participant=unique(eyeY$participant), cue_category = "random")
rand <- cbind(rand,randsample)
colnames(rand)<-colnames(eyeY)
combinedY <- rbind(eyeY,rand)

## perform cbpt
eyeClust <-permuco::clusterlm(combinedY[,(3:ncol(combinedY))] ~ cue_category + Error(participant/cue_category), 
                               data = combinedY[,-(3:ncol(combinedY))],multcomp ="clustermass", 
                               np = 10000)
plot(eyeClust, nbbaselinepts = 65, nbptsperunit = .90)
summary(eyeClust)
sig_LTM <- (which(summary(eyeClust,table_type = "full")$cue_category$`P(>) clustermass`<0.05))*timepersample


#WM
toward_WM <- toward[toward$cue_category=="WM",] %>% dplyr::select(participant,cueFrame, meanyp, cue_category)
eyeY <- toward_WM %>% pivot_wider(names_from = cueFrame, values_from = meanyp)

randsample <- NULL
for (i in 1:length(unique(eyeY$participant))){
  r<-rnorm(n = length(table_tb_t$sd[table_tb_t$cue_category=="WM"]), mean = 0, sd = mean(table_tb_t$sd[table_tb_t$cue_category=="WM"]))
  randsample <- c(randsample,r)
}

randsample <- data.frame(matrix(randsample,nrow = length(unique(eyeY$participant)), ncol = length(table_tb_t$sd[table_tb_t$cue_category=="WM"]),byrow = T))
rand <- data.frame(participant=unique(eyeY$participant), cue_category = "random")
rand <- cbind(rand,randsample)
colnames(rand)<-colnames(eyeY)
combinedY <- rbind(eyeY,rand)

# perform cbpt
eyeClust <-permuco::clusterlm(combinedY[,(3:ncol(combinedY))] ~ cue_category + Error(participant/cue_category), 
                               data = combinedY[,-(3:ncol(combinedY))],multcomp ="clustermass", 
                               np = 10000)
plot(eyeClust, nbbaselinepts = 65, nbptsperunit = .90)
summary(eyeClust)
sig_WM <- (which(summary(eyeClust,table_type = "full")$cue_category$`P(>) clustermass`<0.05))*timepersample

# Comparing two
toward_WM_LTM <- toward %>% dplyr::select(participant,cueFrame, meanyp, cue_category)
eyeY <- toward_WM_LTM %>% pivot_wider(names_from = cueFrame, values_from = meanyp)
eyeY <- eyeY[order(eyeY$cue_category),]

# perform cbpt
eyeClust <-permuco::clusterlm(eyeY[,(3:ncol(eyeY))] ~ cue_category + Error(participant/cue_category), 
                                data = eyeY[,-(3:ncol(eyeY))],multcomp ="clustermass", 
                                np = 10000)
plot(eyeClust, nbbaselinepts = baseline, nbptsperunit = .90)
summary(eyeClust)
sig_LTM_WM <- (which(summary(eyeClust,table_type = "full")$cue_category$`P(>) clustermass`<0.05))*timepersample

(towardness_sigY <-towardnessY + annotate("point", x = sig_LTM_WM , y =-3, colour = "black", size=1) + 
    annotate("point", x = sig_WM , y =-2, colour = rgb(65/255,143/255,111/255),size=1)+
    annotate("point", x = sig_LTM , y =-1, colour = rgb(138/255,118/255,190/255),size=1))
    # scale_y_continuous(breaks = c(0,1,2),limits = c(-1,2))

ggsave("figures/towardness_TB.svg", device = "svg", width = 25, height = 15, units = "cm")
```